---
title: "Site selector data visualization tool"
author: "Molly 'Caz' Bell and A.J. Brown using R version 4.2.2"
date: "`r Sys.Date()`"
output: 
  html_document:
    output_dir: "site_selector_temporary_results"
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
---
# Introduction and Metadata
The following is an interactive data report containing water quality data from 
samples taken by the Colorado State University (CSU) Agriculture Water Quality 
Program (AWQP) at a location selected by the user when running the script
(reported below). For more information, please contact A.J. Brown at
Ansley.Brown@colostate.edu or visit the CSU AWQP Website at 
https://waterquality.colostate.edu.

**Please Note:** All analytes are reported in mg/L *EXCEPT*
selenium (reported in ug/L), specific electrical conductivity (EC; mS/cm), and
pH (log H).

**Misc. Notes**

*	All “0” values reported are non-detects (ND) reported by ALS
*	Total suspended solids (TSS), Electrical Conductivity (EC), and pH are performed by AWQP staff at CSU, whereas all other analytes are reported by ALS Laboratories
*	Methods used to derive each analyte can be found in the Raw Data table in the “method” column

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# set working dir
knitr::opts_knit$set(root.dir = here::here())
   # sets the working directory to where the project file is
```

```{r, include=FALSE}
# load required packages
package.list <- c('plotly',
                  'dplyr',
                  'tidyr',
                  'DT',
                  'ggplot2'
                  # 'gridExtra',
                  # 'GGally',
                  # 'PerformanceAnalytics'
                  )
packageLoad <- function(packages){
  for (i in packages) {
    if (!require(i, character.only = TRUE)) {
      install.packages(i, repos = "http://cran.us.r-project.org")
      library(i, character.only = TRUE)
    }
  }
}
packageLoad(package.list)
```

```{r, include=FALSE}
# Uploading file-merge code to create data frame
source('./Code/file-merger.R')
# create data frame
  # note that directory is defined in file_merger.R file
  # change data file path in that script to modify the data source
dat <- returnAllFiles(d = directory, export = FALSE)
dat2 <- dat
# make columns lowercase so they are all the same case
colnames(dat) <- tolower(names(dat))
```

```{r, include=FALSE}
# Site list
  # To select site, un-comment your location of interest (alphabetical order)
  # This list may need to be updated as more sites are added/removed
# loc = 'ARDEC' #temporarily ARDEC 2200 until I fix file-merger.r
 loc = 'AVRC STAR'
# loc = 'AVRC Cowpea'
# loc = 'Barley'
# loc = 'Berthoud'
# loc = 'Big Hollow'
# loc = 'Boulder Lake'
# loc = 'Below Stagecoach Dam'
# loc = 'Fruita A'
# loc = 'Fruita B'
# loc = 'Fruita NT'
# loc = 'Fruita W'
# loc = 'Fruita C'
# loc = 'Fruita F'
# loc = 'Gunnison'
# loc = 'Kerbel'
# loc = 'Legacy'
# loc = 'Molina'
# loc = 'Stagecoach'
# loc = 'The Ranch'
# loc = 'Upper Yampa'
# loc = 'Yellow Jacket'
# loc = 'North Sand Creek'

#create a filtered df to use for violin plots and scatterplot
filter_dataframe <- function(df, selected_location, include_blanks = FALSE) {
  if (include_blanks) {
    filtered_df <- df %>% filter(location.name == selected_location)
  } else {
    filtered_df <- df %>% filter(location.name == selected_location & event.type != "Lab Blank")
  }
  # make event.type, event.count, location.name, and treatment.name factors
  filtered_df$event.type <- as.factor(filtered_df$event.type)
  filtered_df$event.count <- as.factor(filtered_df$event.count)
  filtered_df$location.name <- as.factor(filtered_df$location.name)
  filtered_df$treatment.name <- as.factor(filtered_df$treatment.name)
  return(filtered_df)
}

# Create df and exclude blanks by default:
filtered_df <- filter_dataframe(dat, loc, include_blanks = FALSE)
```


**Location Selected: <span style="color:#4CAF50">`r loc`</span>**

```{r, include=FALSE}
# function to save site data as csv
csv_data <- function(df, location_name) {
  location_data <- df %>%
    filter(location.name == location_name)
  #create output folder
  folder_name <- "site_selector_temporary_results"
  if (!file.exists(folder_name)) {
    dir.create(folder_name)
  }
  
  file_name <- paste0(folder_name, "/", location_name, "_data_", format(Sys.Date(), "%Y-%m-%d"), ".csv")
  write.csv(location_data, file = file_name, row.names = FALSE)
  return(file_name)
}
# use the function to save
file = csv_data(filtered_df, loc)
```

```{r}
# Creating some subset data for static graphing later
analytes_plot1 <- c("Nitrogen, Nitrate (As N)", 
                    "Nitrogen, Nitrite  (As N)", # note the extra space before "(" due to ALS formatting
                    "Nitrogen, Total Kjeldahl")
analytes_plot2 <- c("Phosphorus, Total (As P)", 
                    "Phosphorus, Total Orthophosphate (as P)",
                    "Suspended Solids (Residue, Non-Filterable)")
analytes_plot3 <- setdiff(unique(filtered_df$analyte), c(analytes_plot1, analytes_plot2))
N_df <- filtered_df %>% filter(analyte %in% analytes_plot1)
P_df <- filtered_df %>% filter(analyte %in% analytes_plot2)
Misc_df <- filtered_df %>% filter(analyte %in% analytes_plot3)
```

**CSV Export of selected site data saved as: **
`r file`

# Events Captured:
```{r}
# Combine event.count and collected values into a single row for each treatment.name and event.count
unique_treatments <- filtered_df %>%
  select(treatment.name, event.count, collected) %>%
  distinct() %>%
  group_by(treatment.name, event.count) %>%
  summarize(
    collected = paste(unique(format(as.Date(collected), "%Y-%m-%d")), collapse = ", "),
    .groups = "drop"  # Explicitly drop the grouping to avoid the warning
  ) %>%
  group_by(treatment.name) %>%
  summarize(
    event.count = paste(unique(event.count), collapse = ", "),
    collected = paste(unique(collected), collapse = ", "),
    .groups = "drop"  # Drop grouping after summarizing
  ) %>%
  arrange(treatment.name)

# Print the table using datatable
datatable(unique_treatments, 
          caption = "Unique Treatments, Event Counts, and Collection Dates (YYYY-MM-DD)")


```

# Analytes Present:
```{r}
# Combine analyte, method, units, mdl, and rl into a unique table
unique_analyte_method <- filtered_df %>%
  select(analyte.abbr, method, units, mdl, rl) %>%
  distinct(analyte.abbr, method, units, mdl, rl) %>%
  arrange(analyte.abbr)

# Print the table using datatable
datatable(unique_analyte_method, 
          caption = "Unique Analytes and Corresponding Methods, Units, MDL, and RL")


```


# Interactive Box and Whisker Plot
Use the below boxplot to graph various water analytes stratified by sample type
(i.e., Inflow, Outflow, etc.). Analytes can be selected by using the dropdown 
menu on the top left of the graph. Hover over points with your mouse for
additional information.

**Instructions**: Use the dropdown menu on the chart to select the analyte of interest.

```{r, warning=FALSE}
# Create the box-and-whisker plot function with hover information
create_boxplot <- function(df, selected_location) {
  
  tryCatch({
    # Get unique analytes for the dropdown menu
    analytes <- unique(df$analyte.abbr)
    
    # Define the default plot (for the first analyte in the list)
    first_analyte <- analytes[1]
    df_filtered <- df[df$analyte.abbr == first_analyte, ]
    
    # Initialize the plot with the first analyte
    box_plot <- plot_ly(data = df_filtered, 
                        x = ~treatment.name, 
                        y = ~result,
                        text = ~paste('Result: ', result,
                                      '<br>Units: ', units, 
                                      '<br>Event Count: ', event.count,
                                      '<br>Date Collected: ', collected),  # Custom hover text
                        hoverinfo = "text",  # Use the custom text for hover info
                        type = 'box',
                        boxpoints = "all",  # Overlay raw data points
                        jitter = 0.3,       # Add jitter to spread out points
                        pointpos = 0,       # Adjust point position
                        colors = "Set2") %>% # Use a color set
      layout(
        title = paste("Box-and-Whisker plots for Location:", selected_location),
        xaxis = list(title = "Treatment Name"),
        yaxis = list(title = "Result (mg/L, ug/L, dS/m, or pH)", range = range(df_filtered$result, na.rm = TRUE)),
        showlegend = TRUE
      )
    
    # Create dropdown buttons for each analyte
    dropdown_buttons <- lapply(analytes, function(analyte_name) {
      df_analyte <- df[df$analyte.abbr == analyte_name, ]
      y_range <- range(df_analyte$result, na.rm = TRUE)  # Calculate y-axis range for each analyte
      
      list(
        method = "update",
        args = list(
          list(
            x = list(df_analyte$treatment.name), 
            y = list(df_analyte$result),
            text = list(paste('Result: ', df_analyte$result,
                              '<br>Units: ', df_analyte$units, 
                              '<br>Event Count: ', df_analyte$event.count,
                              '<br>Date Collected: ', df_analyte$collected))
          ),
          list(
            yaxis = list(range = c(y_range[1]-0.05*y_range[2], y_range[2]+0.05*y_range[2]))
          )
        ),
        label = analyte_name
      )
    })
    
    # Add the dropdown menu to the plot
    box_plot <- box_plot %>%
      layout(
        updatemenus = list(
          list(
            buttons = dropdown_buttons,
            direction = "down",
            showactive = TRUE,
            x = 0.1, y = 1.1,
            xanchor = 'left', yanchor = 'top'
          )
        )
      )
    
    return(box_plot)
    
  }, error = function(e) {
    cat("An error occurred:", conditionMessage(e), "\n")
    return(NULL)
  })
}

# Example call to the function
create_boxplot(filtered_df, loc)

```

**To save your plot as a PNG:** hover over the top right corner and click
the camera icon.  The PNG will be saved in your downloads folder.


# Scatterplot Matrix of All Numerical Analyte Data
Use the below scatter matrix to look at potential relationships between
numerical variables. Black points represent Inflow samples, and red points
represent Outflow samples

```{r, warning=FALSE}
create_scatterplot_matrix <- function(df, location_name) {
  
  tryCatch({
    #special filtering for scatterplot matrix, different from above
    filtered_data <- df %>%
      mutate(sample.id = gsub("-[1-5]$", "", sample.id), # remove -1 to -5 at the end of strings 
             sample.id = gsub("-[1-5]-D$", "-D", sample.id),  # do the same for duplicates
             sample.id = gsub("-2,3,5", "", sample.id)) %>%  # remove special case of "-2,3,5"
      select(sample.id, result, analyte.abbr, event.type)

    # Pivot the data wider with aggregation of duplicates by mean
    pivoted_data <- filtered_data %>%
      tidyr::pivot_wider(
        names_from = analyte.abbr,
        values_from = result,
        values_fn = list(result = mean)  # Specify how to aggregate duplicates
      )

    # raise error if no data provided
    if (nrow(pivoted_data) == 0) {
      stop("No data available for the given location.")
    }

    # Exclude the first two columns (sample.id and event.type)
    analyte_columns <- colnames(pivoted_data)[-c(1, 2)]  

    # Create scatterplot matrix using pairs()
    pairs(pivoted_data[, analyte_columns],
          labels = analyte_columns,
          col = as.numeric(factor(pivoted_data$event.type)),
          pch = 19)

    # TODO: create scatter matrix using PerformanceAnalytics package
    # chart.Correlation(pd2,
    #                 histogram = T,
    #                 method = 'pearson',
    #                 pch = 8
    #                 )
  }, error = function(e) {
    cat("An error occurred when producing the scatter matrix:",
        conditionMessage(e))
  })
}

create_scatterplot_matrix(filtered_df, loc)
```

```{r, fig.height=20, fig.width=12}
# Create a check for treatment.name levels, aside from "Inflow", and make treatment.name the x-axis if present (e.g., 'CT1').
if (length(levels(as.factor(filtered_df$treatment.name))) > 1) {
  x_axis_var_name <- "treatment.name"
} else {
  x_axis_var_name <- "event.type"
}


```

```{r, eval=FALSE}
create_and_save_boxwhisker_plot <- function(df, output_dir, file_name, plot.title, x_axis_var) {
  tryCatch({
    # Ensure the output directory exists
    if (!dir.exists(output_dir)) {
      dir.create(output_dir, recursive = TRUE)
    }
    
    # Create the file paths for portrait and landscape versions
    file_path_portrait <- file.path(output_dir, file_name)
    file_path_landscape <- file.path(output_dir, paste0(tools::file_path_sans_ext(file_name), "_landscape.jpg"))
    
    # Generate the box and whisker plot with dynamic x-axis
    bw_plot <- ggplot(df, aes_string(x = x_axis_var, y = "result", fill = x_axis_var)) +
      geom_boxplot(position = position_dodge(0)) + 
      geom_jitter(width = 0.2, size = 1.5, alpha = 0.6) + # Add jitter to show individual data points
      facet_grid(analyte.abbr ~ event.count, scales = "free_y", space = "free_x") +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14), # Increase x axis text size
        axis.text.y = element_text(size = 14), # Increase y axis text size
        axis.title = element_text(size = 16), # Increase axis title text size
        strip.text = element_text(size = 16), # Increase facet strip text size
        legend.text = element_text(size = 14), # Increase legend text size
        legend.title = element_text(size = 16), # Increase legend title size
        plot.title = element_text(size = 20, hjust = 0.5), # Increase plot title size
        plot.subtitle = element_text(size = 18), # Increase plot subtitle size
        plot.caption = element_text(size = 12), # Increase plot caption size
        strip.background = element_rect(colour = "black", fill = "white"),
        panel.grid.major = element_line(size = 1), # Increase major grid line size
        panel.grid.minor = element_line(size = 0.5) # Increase minor grid line size
      ) +
      labs(title = plot.title, x = "Event Type", y = "Result", fill = "Event Type") +
      scale_fill_brewer(palette = "Pastel1") # Colorblind-friendly palette
    
    # Save the portrait version of the plot
    ggsave(file_path_portrait, bw_plot, width = 8.5, height = 11, units = "in")
    
    # Save the landscape version of the plot
    ggsave(file_path_landscape, bw_plot, width = 11, height = 8.5, units = "in")
    
    # Return the plot object for rendering in the HTML document
    return(bw_plot)
  }, error = function(e) {
    message("An error occurred in create_and_save_boxwhisker_plot: ", e$message)
    return(e$message) # Return error message
  })
}



bw_N <- create_and_save_boxwhisker_plot(N_df,
                                        output_dir = "./site_selector_temporary_results",
                                        file_name = "bw_N.jpg",
                                        plot.title = 'Forms of Nitrogen',
                                        x_axis_var = x_axis_var_name)
bw_P <- create_and_save_boxwhisker_plot(P_df,
                                        output_dir = "./site_selector_temporary_results",
                                        file_name = "bw_P.jpg",
                                        plot.title = 'Forms of Phosphorous',
                                        x_axis_var = x_axis_var_name)
bw_Misc <- create_and_save_boxwhisker_plot(Misc_df,
                                           output_dir = "./site_selector_temporary_results",
                                           file_name = "bw_Misc.jpg",
                                           plot.title = 'Water Quality Analytes',
                                           x_axis_var = x_axis_var_name)
```

# Static Bar Plots
```{r}
# Static Bar Plots with Raw Data Points Overlay
create_and_save_bar_plot <- function(df, output_dir, file_name, plot.title, x_axis_var) {
  tryCatch({
    # Ensure the output directory exists
    if (!dir.exists(output_dir)) {
      dir.create(output_dir, recursive = TRUE)
    }
    
    # Summarize the data for the bar plot
    summary_df <- df %>%
      group_by(!!sym(x_axis_var), analyte.abbr, event.count) %>%
      summarize(mean = mean(result, na.rm = TRUE),
                sd = sd(result, na.rm = TRUE),
                .groups = 'drop')
    
    # Ensure "Inflow" is the first level if present
    summary_df[[x_axis_var]] <- factor(summary_df[[x_axis_var]], levels = c("Inflow", setdiff(unique(summary_df[[x_axis_var]]), "Inflow")))

    # Create the file paths for portrait and landscape versions
    file_path_portrait <- file.path(output_dir, file_name)
    file_path_landscape <- file.path(output_dir, paste0(tools::file_path_sans_ext(file_name), "_landscape.jpg"))
    
    # Generate the bar plot with error bars and raw data points
    bar_plot <- ggplot(summary_df, aes(x = !!sym(x_axis_var), y = mean, fill = !!sym(x_axis_var))) +
      geom_bar(stat = "identity", position = position_dodge(0), color = 'black') +
      geom_errorbar(aes(ymin = pmax(mean - sd, 0), ymax = mean + sd), width = 0.2, position = position_dodge(0)) +
      # Add raw data points using geom_jitter
      geom_jitter(data = df, aes(x = !!sym(x_axis_var), y = result), 
                  width = 0.2, alpha = 0.6, size = 1, show.legend = FALSE) +  # Jitter to avoid overlap
      facet_grid(analyte.abbr ~ event.count, scales = "free_y", space = "free_x") +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14), # Increase x-axis text size
        axis.text.y = element_text(size = 14), # Increase y-axis text size
        axis.title = element_text(size = 16), # Increase axis title text size
        strip.text = element_text(size = 16), # Increase facet strip text size
        legend.text = element_text(size = 14), # Increase legend text size
        legend.title = element_text(size = 16), # Increase legend title size
        plot.title = element_text(size = 20, hjust = 0.5), # Increase plot title size
        plot.subtitle = element_text(size = 18), # Increase plot subtitle size
        plot.caption = element_text(size = 12), # Increase plot caption size
        strip.background = element_rect(colour = "black", fill = "white"),
        panel.grid.major = element_line(linewidth = 1),
        panel.grid.minor = element_line(linewidth = 0.5)
      ) +
      labs(title = plot.title, x = "Event Type", y = "Result", fill = "Event Type") +
      scale_fill_brewer(palette = "Pastel1") # Colorblind-friendly palette

    # Save the portrait version of the plot
    ggsave(file_path_portrait, bar_plot, width = 8.5, height = 11, units = "in")
    
    # Save the landscape version of the plot
    ggsave(file_path_landscape, bar_plot, width = 11, height = 8.5, units = "in")

    # Return the plot object for rendering in the HTML document
    return(bar_plot)
  }, error = function(e) {
    message("An error occurred in create_and_save_bar_plot: ", e$message)
    return(e$message) # Return error message
  })
}
```

## Forms of N
```{r}
bar_N <- create_and_save_bar_plot(N_df,
                                  output_dir = "./site_selector_temporary_results",
                                  file_name = "bar_N.jpg",
                                  plot.title = 'Forms of Nitrogen',
                                  x_axis_var = x_axis_var_name)
bar_N
```

## Forms of P and TSS
```{r}
bar_P <- create_and_save_bar_plot(P_df,
                                  output_dir = "./site_selector_temporary_results",
                                  file_name = "bar_P.jpg",
                                  plot.title = 'Forms of Phosphorus',
                                  x_axis_var = x_axis_var_name)
bar_P
```

## Other miscellaneous analytes
```{r}
bar_Misc <- create_and_save_bar_plot(Misc_df,
                                     output_dir = "./site_selector_temporary_results",
                                     file_name = "bar_Misc.jpg",
                                     plot.title = 'Water Quality Analytes',
                                     x_axis_var = x_axis_var_name)
bar_Misc
```

## Annual means for each analyte
```{r}
# Updated Faceted Bar Plot by Analyte with Units, Matching Colors, Y-axis Min at 0, and "Inflow" Ordered First
# Updated Faceted Bar Plot with Optional Stratification by Method
create_facet_bar_plot <- function(df, output_dir, file_name, x_axis_var, stratify_by_method = FALSE) {
  tryCatch({
    # Ensure the output directory exists
    if (!dir.exists(output_dir)) {
      dir.create(output_dir, recursive = TRUE)
    }

    # Extract unique analytes and their corresponding units
    analyte_units <- df %>% 
      select(analyte.abbr, units) %>% 
      distinct()

    # Summarize data and merge units
    summary_df <- df %>%
      group_by(!!sym(x_axis_var), analyte.abbr, method.name) %>%
      summarize(mean = mean(result, na.rm = TRUE),
                sd = sd(result, na.rm = TRUE),
                .groups = 'drop') %>%
      left_join(analyte_units, by = "analyte.abbr")  # Join to get the units

    # Ensure "Inflow" is the first level if present
    summary_df[[x_axis_var]] <- factor(summary_df[[x_axis_var]], levels = c("Inflow", setdiff(unique(summary_df[[x_axis_var]]), "Inflow")))

    # Create custom facet labels with analyte and unit, ensuring a closing parenthesis
    facet_labels <- summary_df %>%
      mutate(facet_label = paste(analyte.abbr, units, sep = " ("), facet_label = paste0(facet_label, ")")) %>%
      distinct(analyte.abbr, facet_label) %>%
      deframe()

    # File path for saving the plot
    file_path <- file.path(output_dir, file_name)

    # Conditionally stratify by method.name if stratify_by_method is TRUE
    if (stratify_by_method) {
      # If stratifying by method, ensure method.name is used in the plot and adjust dodge width
      facet_bar_plot <- ggplot(summary_df, aes(x = !!sym(x_axis_var), y = mean, fill = method.name)) +
        geom_bar(stat = "identity", position = position_dodge(width = 0.7), color = "black") +
        geom_errorbar(aes(ymin = pmax(mean - sd, 0), ymax = mean + sd), width = 0.2, position = position_dodge(width = 0.7))
    } else {
      # When not stratifying by method, remove method.name grouping to prevent redundant bars
      summary_df <- summary_df %>%
        group_by(!!sym(x_axis_var), analyte.abbr) %>%
        summarize(mean = mean(mean), sd = mean(sd), .groups = "drop")

      facet_bar_plot <- ggplot(summary_df, aes(x = !!sym(x_axis_var), y = mean, fill = !!sym(x_axis_var))) +
        geom_bar(stat = "identity", position = "dodge", color = "black") +
        geom_errorbar(aes(ymin = pmax(mean - sd, 0), ymax = mean + sd), width = 0.2, position = position_dodge(width = 0.9)) +
        geom_jitter(data = df, aes(x = !!sym(x_axis_var), y = result), 
                    width = 0.2, alpha = 0.6, size = 1, show.legend = FALSE)
    }

    # Add facets, theme, labels, and save the plot
    facet_bar_plot <- facet_bar_plot +
      facet_wrap(~ analyte.abbr, scales = "free_y", labeller = as_labeller(facet_labels)) +  # Use custom facet labels
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 10),
        legend.position = "bottom"
      ) +
      labs(title = "Average Result by Analyte",
           x = "Event Type",
           y = "Mean Result",
           fill = ifelse(stratify_by_method, "Method", x_axis_var)) +
      scale_fill_brewer(palette = "Pastel1") +  # Adjust this to match your preferred palette
      coord_cartesian(ylim = c(0, NA))  # Ensure y-axis starts at 0

    # Save the plot
    ggsave(file_path, facet_bar_plot, width = 10, height = 8, units = "in")

    # Return the plot object for rendering in the HTML document
    return(facet_bar_plot)
  }, error = function(e) {
    message("An error occurred in create_facet_bar_plot: ", e$message)
    return(NULL)
  })
}
```

### By Method and Treatment
```{r}
# Face with stratification by method
facet_bar_with_method <- create_facet_bar_plot(
  filtered_df,
  output_dir = "./site_selector_temporary_results",
  file_name = "facet_bar_analytes_method_annual.jpg",
  x_axis_var = x_axis_var_name,
  stratify_by_method = TRUE
)
facet_bar_with_method
```

### By Treatment Only
```{r}
# Facet bar without method
facet_bar_without_method <- create_facet_bar_plot(
  filtered_df,
  output_dir = "./site_selector_temporary_results",
  file_name = "facet_bar_analytes_annual.jpg",
  x_axis_var = x_axis_var_name,
  stratify_by_method = FALSE
)
facet_bar_without_method
```


# Raw Data
Use the table below to look at raw values.
```{r}
datatable(filtered_df)
```