---
title: "Site selector data visualization tool"
author: "Molly 'Caz' Bell and A.J. Brown using R version 4.2.2"
date: "`r Sys.Date()`"
output: 
  html_document:
    output_dir: "site_selector_temporary_results"
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
---

# Introduction and Metadata
The following is an interactive data report containing water quality data from 
samples taken by the Colorado State University (CSU) Agriculture Water Quality 
Program (AWQP) at a location selected by the user when running the script
(reported below). For more information, please contact A.J. Brown at
Ansley.Brown@colostate.edu or visit the CSU AWQP Website at 
https://waterquality.colostate.edu.

**Please Note:** All analytes are reported in mg/L *EXCEPT*
selenium (reported in ug/L), specific electrical conductivity (EC; mS/cm), and
pH (log H).

**Misc. Notes**

*	All “0” values reported are non-detects (ND) reported by ALS
*	Total suspended solids (TSS), Electrical Conductivity (EC), and pH are performed by AWQP staff at CSU, whereas all other analytes are reported by ALS Laboratories
*	Methods used to derive each analyte can be found in the Raw Data table in the “method” column

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}
# packages
package.list <- c('plotly','dplyr','tidyr','DT','ggplot2','lubridate','tools','stringr','tibble')
packageLoad <- function(packages){
  for (i in packages) {
    if (!require(i, character.only = TRUE)) {
      install.packages(i, repos = "http://cran.us.r-project.org")
      library(i, character.only = TRUE)
    }
  }
}
packageLoad(package.list)
```

```{r, include=FALSE}
# Data build
source('./Code/file-merger.R')
dat <- returnAllFiles(d = directory, export = FALSE)
colnames(dat) <- tolower(names(dat))
```

```{r, include=FALSE}
# ---- Site list (toggle exactly ONE 'loc' line) ----
# loc <- 'ARDEC'                     
# loc <- 'AVRC STAR'
# loc <- 'AVRC Cowpea'
# loc <- 'Barley'
# loc <- 'Berthoud'
# loc <- 'Big Hollow'
# loc <- 'Boulder Lake'
# loc <- 'Below Stagecoach Dam'
# loc <- 'Fruita A'
# loc <- 'Fruita B'
# loc <- 'Fruita NT'
# loc <- 'Fruita W'
# loc <- 'Fruita C'
# loc <- 'Fruita F'
# loc <- 'Gunnison'
loc <- 'Kerbel'                      
# loc <- 'Legacy'
# loc <- 'Molina'
# loc <- 'Stagecoach'
# loc <- 'The Ranch'
# loc <- 'Upper Yampa'
# loc <- 'Yellow Jacket'
# loc <- 'North Sand Creek'

all_sites <- c(
  'ARDEC','AVRC STAR','AVRC Cowpea','Barley','Berthoud','Big Hollow','Boulder Lake',
  'Below Stagecoach Dam','Fruita A','Fruita B','Fruita NT','Fruita W','Fruita C','Fruita F',
  'Gunnison','Kerbel','Legacy','Molina','Stagecoach','The Ranch','Upper Yampa',
  'Yellow Jacket','North Sand Creek'
)
if (!exists("loc")) stop("Please set 'loc' to one of the known sites.")
if (!loc %in% all_sites) stop(paste("Unknown 'loc':", loc))
```

```{r, include=FALSE}
# ---- TZ normalization helpers ----
normalize_to_denver <- function(x){
  if (inherits(x, "POSIXct")) {
    tzx <- attr(x, "tzone")
    if (is.null(tzx) || tzx == "") {
      return(lubridate::force_tz(x, "America/Denver"))
    } else {
      return(lubridate::with_tz(x, "America/Denver"))
    }
  } else if (is.character(x)) {
    parsed <- suppressWarnings(lubridate::parse_date_time(
      x, orders = c("Y-m-d H:M:S","Y-m-d H:M","m/d/Y H:M:S","m/d/Y H:M","Y-m-d"),
      tz = "America/Denver"
    ))
    return(lubridate::force_tz(parsed, "America/Denver"))
  } else {
    return(x)
  }
}

filter_dataframe <- function(df, selected_location, include_blanks = FALSE) {
  if (include_blanks) {
    filtered_df <- df %>% dplyr::filter(location.name == selected_location)
  } else {
    filtered_df <- df %>% dplyr::filter(location.name == selected_location & event.type != "Lab Blank")
  }
  filtered_df$event.type     <- as.factor(filtered_df$event.type)
  filtered_df$event.count    <- as.factor(filtered_df$event.count)
  filtered_df$location.name  <- as.factor(filtered_df$location.name)
  filtered_df$treatment.name <- as.factor(filtered_df$treatment.name)
  filtered_df <- filtered_df %>% dplyr::mutate(collected_local = normalize_to_denver(collected))
  filtered_df$collected_disp <- format(filtered_df$collected_local, "%m/%d/%Y %H:%M")
  filtered_df
}

filtered_df <- filter_dataframe(dat, loc, include_blanks = FALSE)

# Canonical event date (one per event), facet label
event_lookup <- filtered_df %>%
  dplyr::mutate(collected_date = as.Date(collected_local)) %>%
  dplyr::group_by(event.count) %>%
  dplyr::summarise(n_dates = dplyr::n_distinct(collected_date),
                   event_date = min(collected_date, na.rm = TRUE),
                   .groups = "drop")
filtered_df <- filtered_df %>% dplyr::left_join(event_lookup %>% dplyr::select(event.count, event_date), by = "event.count")
filtered_df <- filtered_df %>% dplyr::mutate(
  event_facet = factor(event.count, levels = unique(event.count[order(event_date)]))
)
label_map <- filtered_df %>% dplyr::select(event_facet, event_date) %>% dplyr::distinct()
event_labeller <- function(values) setNames(paste0(values, "\n", format(label_map$event_date[match(values, label_map$event_facet)], "%m-%d")), values)
```


**Location Selected: <span style="color:#4CAF50">`r loc`</span>**

```{r, include=FALSE}
# Export selected-site CSV
csv_data <- function(df, location_name) {
  location_data <- df %>% dplyr::filter(location.name == location_name)
  folder_name <- "site_selector_temporary_results"
  if (!file.exists(folder_name)) dir.create(folder_name)
  file_name <- paste0(folder_name, "/", location_name, "_data_", format(Sys.Date(), "%Y-%m-%d"), ".csv")
  write.csv(location_data, file = file_name, row.names = FALSE)
  return(file_name)
}
file = csv_data(filtered_df, loc)
```

```{r}
# Subsets
analytes_plot1 <- c("Nitrogen, Nitrate (As N)", "Nitrogen, Nitrite  (As N)", "Nitrogen, Total Kjeldahl")
analytes_plot2 <- c("Phosphorus, Total (As P)", "Phosphorus, Total Orthophosphate (as P)", "Suspended Solids (Residue, Non-Filterable)")
analytes_plot3 <- setdiff(unique(filtered_df$analyte), c(analytes_plot1, analytes_plot2))

N_df    <- filtered_df %>% dplyr::filter(analyte %in% analytes_plot1)
P_df    <- filtered_df %>% dplyr::filter(analyte %in% analytes_plot2)
Misc_df <- filtered_df %>% dplyr::filter(analyte %in% analytes_plot3)
```

**CSV Export of selected site data saved as:** `r file`

# Events Captured
```{r}
unique_treatments <- filtered_df %>%
  dplyr::select(treatment.name, event.count, event_date) %>%
  dplyr::distinct() %>%
  dplyr::group_by(treatment.name, event.count) %>%
  dplyr::summarise(
    collected = paste(sort(unique(format(event_date, "%Y-%m-%d"))), collapse = ", "),
    .groups = "drop"
  ) %>%
  dplyr::arrange(treatment.name, event.count)

DT::datatable(unique_treatments, caption = "Unique Treatments, Event Counts, and Collection Dates (YYYY-MM-DD)")
```

# Interactive Box & Whisker Plot
```{r, warning=FALSE}
create_boxplot <- function(df, selected_location) {
  tryCatch({
    analytes <- unique(df$analyte.abbr)
    first_analyte <- analytes[1]
    df_filtered <- df[df$analyte.abbr == first_analyte, ]
    p <- ggplot(df_filtered, aes(x = treatment.name, y = as.numeric(result), fill = treatment.name)) +
      geom_boxplot(position = position_dodge(0), na.rm = TRUE) +
      geom_jitter(width = 0.2, size = 1.5, alpha = 0.6, na.rm = TRUE) +
      facet_grid(analyte.abbr ~ event_facet, scales = "free_y", space = "free_x",
                 labeller = labeller(event_facet = event_labeller)) +
      theme_minimal() +
      labs(title = paste("Box-and-Whisker plots for Location:", selected_location),
           x = "Event Type", y = "Result", fill = "Event Type")
    ggplotly(p, tooltip = c("y"))
  }, error = function(e) {
    cat("An error occurred:", conditionMessage(e), "\n"); NULL
  })
}
create_boxplot(filtered_df, loc)
```

```{r, fig.height=20, fig.width=12}
# dynamic x-axis in static plots
if (length(levels(as.factor(filtered_df$treatment.name))) > 1) {
  x_axis_var_name <- "treatment.name"
} else {
  x_axis_var_name <- "event.type"
}
```

# Static Bar Plots (Per Event) — with Error Bars
```{r}
library(ggplot2); library(dplyr); library(tools)

create_and_save_bar_plot_event <- function(df, output_dir, file_name, plot.title, x_axis_var) {
  tryCatch({
    if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
    df <- df %>% dplyr::mutate(result = suppressWarnings(as.numeric(result))) %>% dplyr::filter(is.finite(result))

    # summary stats per event facet, analyte, and x-axis (treatment/event.type)
    summary_df <- df %>%
      dplyr::group_by(.data[[x_axis_var]], analyte.abbr, event_facet) %>%
      dplyr::summarise(mean = mean(result, na.rm = TRUE),
                       sd = sd(result, na.rm = TRUE),
                       n = dplyr::n(),
                       .groups = "drop")

    # facet ordering & labels
    summary_df$event_facet <- factor(summary_df$event_facet, levels = levels(df$event_facet))

    file_path_portrait  <- file.path(output_dir, file_name)
    file_path_landscape <- file.path(output_dir, paste0(tools::file_path_sans_ext(file_name), "_landscape.jpg"))

    bar_plot <- ggplot(summary_df, aes(x = .data[[x_axis_var]], y = mean, fill = .data[[x_axis_var]])) +
      geom_bar(stat = "identity", position = position_dodge(0), color = "black") +
      geom_errorbar(aes(ymin = pmax(mean - sd, 0), ymax = mean + sd),
                    width = 0.2, position = position_dodge(0)) +
      # raw points (jitter) in background
      geom_jitter(data = df, aes(x = .data[[x_axis_var]], y = result), width = 0.2, alpha = 0.6,
                  size = 1, show.legend = FALSE) +
      facet_grid(analyte.abbr ~ event_facet, scales = "free_y", space = "free_x",
                 labeller = labeller(event_facet = event_labeller)) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
            strip.text  = element_text(size = 12),
            legend.position = "bottom") +
      labs(title = plot.title, x = "Event Type", y = "Mean Result", fill = "Event Type") +
      scale_fill_brewer(palette = "Pastel1")

    ggsave(file_path_portrait,  bar_plot, width = 8.5, height = 11, units = "in")
    ggsave(file_path_landscape, bar_plot, width = 11, height = 8.5, units = "in")
    bar_plot
  }, error = function(e) { message("Error in create_and_save_bar_plot_event: ", e$message); NULL })
}

# Calls
bar_N    <- create_and_save_bar_plot_event(N_df, "./site_selector_temporary_results", "bar_N.jpg", "Forms of Nitrogen (per Event)", x_axis_var_name); bar_N
bar_P    <- create_and_save_bar_plot_event(P_df, "./site_selector_temporary_results", "bar_P.jpg", "Forms of Phosphorus (per Event)", x_axis_var_name); bar_P
bar_Misc <- create_and_save_bar_plot_event(Misc_df, "./site_selector_temporary_results", "bar_Misc.jpg", "Water Quality Analytes (per Event)", x_axis_var_name); bar_Misc
```

# Annual Means by Analyte & Treatment (Error Bars)
```{r}
create_facet_bar_plot_annual <- function(df, output_dir, file_name, x_axis_var, stratify_by_method = FALSE) {
  tryCatch({
    if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)
    df <- df %>% dplyr::mutate(result = suppressWarnings(as.numeric(result))) %>% dplyr::filter(is.finite(result))

    # Ensure "Inflow" (if present) is first in x-axis
    if (x_axis_var %in% names(df)) {
      df[[x_axis_var]] <- factor(df[[x_axis_var]], levels = c("Inflow", setdiff(unique(df[[x_axis_var]]), "Inflow")))
    }

    analyte_units <- df %>% dplyr::select(analyte.abbr, units) %>% dplyr::distinct()

    if (stratify_by_method) {
      summary_df <- df %>%
        dplyr::group_by(.data[[x_axis_var]], analyte.abbr, method.name) %>%
        dplyr::summarise(mean = mean(result, na.rm = TRUE), sd = sd(result, na.rm = TRUE), n = dplyr::n(), .groups = "drop")
      p <- ggplot(summary_df, aes(x = .data[[x_axis_var]], y = mean, fill = method.name)) +
        geom_bar(stat = "identity", position = position_dodge(width = 0.7), color = "black") +
        geom_errorbar(aes(ymin = pmax(mean - sd, 0), ymax = mean + sd),
                      width = 0.2, position = position_dodge(width = 0.7))
    } else {
      summary_df <- df %>%
        dplyr::group_by(.data[[x_axis_var]], analyte.abbr) %>%
        dplyr::summarise(mean = mean(result, na.rm = TRUE), sd = sd(result, na.rm = TRUE), n = dplyr::n(), .groups = "drop")
      p <- ggplot(summary_df, aes(x = .data[[x_axis_var]], y = mean, fill = .data[[x_axis_var]])) +
        geom_bar(stat = "identity", position = "dodge", color = "black") +
        geom_errorbar(aes(ymin = pmax(mean - sd, 0), ymax = mean + sd),
                      width = 0.2, position = position_dodge(width = 0.9))
    }

    # facet labels include units
    facet_labels <- analyte_units %>%
      dplyr::mutate(lab = paste0(analyte.abbr, " (", units, ")")) %>%
      tibble::deframe()

    p <- p +
      facet_wrap(~ analyte.abbr, scales = "free_y", labeller = as_labeller(facet_labels)) +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1),
            strip.text = element_text(size = 10),
            legend.position = "bottom") +
      labs(title = ifelse(stratify_by_method, "Annual Mean by Analyte (Stratified by Method)", "Annual Mean by Analyte"),
           x = "Event Type", y = "Mean Result",
           fill = ifelse(stratify_by_method, "Method", x_axis_var)) +
      scale_fill_brewer(palette = "Pastel1") +
      coord_cartesian(ylim = c(0, NA))

    file_path <- file.path(output_dir, file_name)
    ggsave(file_path, p, width = 10, height = 8, units = "in")
    p
  }, error = function(e) { message("Error in create_facet_bar_plot_annual: ", e$message); NULL })
}

facet_bar_without_method <- create_facet_bar_plot_annual(filtered_df, "./site_selector_temporary_results", "facet_bar_analytes_annual.jpg", x_axis_var_name, stratify_by_method = FALSE); facet_bar_without_method
facet_bar_with_method    <- create_facet_bar_plot_annual(filtered_df, "./site_selector_temporary_results", "facet_bar_analytes_method_annual.jpg", x_axis_var_name, stratify_by_method = TRUE); facet_bar_with_method
```

# Raw Data
```{r}
DT::datatable(
  filtered_df %>%
    dplyr::mutate(collected = collected_disp) %>%
    dplyr::select(-collected_disp, -collected_local)
)
```
