---
title: "Site selector data visualization tool"
author: "Molly 'Caz' Bell and A.J. Brown using R version 4.2.2"
date: "`r Sys.Date()`"
output: 
  html_document:
    output_dir: "site_selector_temporary_results"
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
---

# Introduction and Metadata
The following is an interactive data report containing water quality data from 
samples taken by the Colorado State University (CSU) Agriculture Water Quality 
Program (AWQP) at a location selected by the user when running the script
(reported below). For more information, please contact A.J. Brown at
Ansley.Brown@colostate.edu or visit the CSU AWQP Website at 
https://waterquality.colostate.edu.

**Please Note:** All analytes are reported in mg/L *EXCEPT*
selenium (reported in ug/L), specific electrical conductivity (EC; mS/cm), and
pH (log H).

**Misc. Notes**

*	All “0” values reported are non-detects (ND) reported by ALS
*	Total suspended solids (TSS), Electrical Conductivity (EC), and pH are performed by AWQP staff at CSU, whereas all other analytes are reported by ALS Laboratories
*	Methods used to derive each analyte can be found in the Raw Data table in the “method” column

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r, include=FALSE}
# packages
package.list <- c('plotly','dplyr','tidyr','DT','ggplot2','lubridate','tools','stringr','tibble', 'tools')
packageLoad <- function(packages){
  for (i in packages) {
    if (!require(i, character.only = TRUE)) {
      install.packages(i, repos = "http://cran.us.r-project.org")
      library(i, character.only = TRUE)
    }
  }
}
packageLoad(package.list)
```

```{r, include=FALSE}
# Data build
source('./Code/file-merger.R')
dat <- returnAllFiles(d = directory, export = FALSE)
colnames(dat) <- tolower(names(dat))
```

```{r, include=FALSE}
# ---- Site list (toggle exactly ONE 'loc' line) ----
# loc <- 'ARDEC'                     
# loc <- 'AVRC STAR'
# loc <- 'AVRC Cowpea'
# loc <- 'Barley'
# loc <- 'Berthoud'
# loc <- 'Big Hollow'
# loc <- 'Boulder Lake'
# loc <- 'Below Stagecoach Dam'
# loc <- 'Fruita A'
# loc <- 'Fruita B'
# loc <- 'Fruita NT'
# loc <- 'Fruita W'
# loc <- 'Fruita C'
# loc <- 'Fruita F'
# loc <- 'Gunnison'
loc <- 'Kerbel'                      
# loc <- 'Legacy'
# loc <- 'Molina'
# loc <- 'Stagecoach'
# loc <- 'The Ranch'
# loc <- 'Upper Yampa'
# loc <- 'Yellow Jacket'
# loc <- 'North Sand Creek'

all_sites <- c(
  'ARDEC','AVRC STAR','AVRC Cowpea','Barley','Berthoud','Big Hollow','Boulder Lake',
  'Below Stagecoach Dam','Fruita A','Fruita B','Fruita NT','Fruita W','Fruita C','Fruita F',
  'Gunnison','Kerbel','Legacy','Molina','Stagecoach','The Ranch','Upper Yampa',
  'Yellow Jacket','North Sand Creek'
)
if (!exists("loc")) stop("Please set 'loc' to one of the known sites.")
if (!loc %in% all_sites) stop(paste("Unknown 'loc':", loc))
```

```{r, include=FALSE}
# ---- TZ normalization helpers ----
normalize_to_denver <- function(x){
  if (inherits(x, "POSIXct")) {
    tzx <- attr(x, "tzone")
    if (is.null(tzx) || tzx == "") {
      return(lubridate::force_tz(x, "America/Denver"))
    } else {
      return(lubridate::with_tz(x, "America/Denver"))
    }
  } else if (is.character(x)) {
    parsed <- suppressWarnings(lubridate::parse_date_time(
      x, orders = c("Y-m-d H:M:S","Y-m-d H:M","m/d/Y H:M:S","m/d/Y H:M","Y-m-d"),
      tz = "America/Denver"
    ))
    return(lubridate::force_tz(parsed, "America/Denver"))
  } else {
    return(x)
  }
}

filter_dataframe <- function(df, selected_location, include_blanks = FALSE) {
  if (include_blanks) {
    filtered_df <- df %>% dplyr::filter(location.name == selected_location)
  } else {
    filtered_df <- df %>% dplyr::filter(location.name == selected_location & event.type != "Lab Blank")
  }
  filtered_df$event.type     <- as.factor(filtered_df$event.type)
  filtered_df$event.count    <- as.factor(filtered_df$event.count)
  filtered_df$location.name  <- as.factor(filtered_df$location.name)
  filtered_df$treatment.name <- as.factor(filtered_df$treatment.name)
  filtered_df <- filtered_df %>% dplyr::mutate(collected_local = normalize_to_denver(collected))
  filtered_df$collected_disp <- format(filtered_df$collected_local, "%m/%d/%Y %H:%M")
  filtered_df
}

filtered_df <- filter_dataframe(dat, loc, include_blanks = FALSE)

# Canonical event date (one per event), facet label
event_lookup <- filtered_df %>%
  dplyr::mutate(collected_date = as.Date(collected_local)) %>%
  dplyr::group_by(event.count) %>%
  dplyr::summarise(n_dates = dplyr::n_distinct(collected_date),
                   event_date = min(collected_date, na.rm = TRUE),
                   .groups = "drop")
filtered_df <- filtered_df %>% dplyr::left_join(event_lookup %>% dplyr::select(event.count, event_date), by = "event.count")
filtered_df <- filtered_df %>% dplyr::mutate(
  event_facet = factor(event.count, levels = unique(event.count[order(event_date)]))
)
label_map <- filtered_df %>% dplyr::select(event_facet, event_date) %>% dplyr::distinct()
event_date_only_labeller <- function(values) {
  dates <- label_map$event_date[match(values, label_map$event_facet)]
  setNames(paste0("Event\nDate\n", format(dates, "%m-%d")), values)
}

event_labeller <- function(values) setNames(paste0(values, "\n", format(label_map$event_date[match(values, label_map$event_facet)], "%m-%d")), values)
```


# Location Selected: <span style="color:#4CAF50">`r loc`</span>

```{r, include=FALSE}
# Export selected-site CSV
csv_data <- function(df, location_name) {
  location_data <- df %>% dplyr::filter(location.name == location_name)
  folder_name <- "site_selector_temporary_results"
  if (!file.exists(folder_name)) dir.create(folder_name)
  file_name <- paste0(folder_name, "/", location_name, "_data_", format(Sys.Date(), "%Y-%m-%d"), ".csv")
  write.csv(location_data, file = file_name, row.names = FALSE)
  return(file_name)
}
file = csv_data(filtered_df, loc)
```

```{r}
# Subsets
analytes_plot1 <- c("Nitrogen, Nitrate (As N)", "Nitrogen, Nitrite  (As N)", "Nitrogen, Total Kjeldahl")
analytes_plot2 <- c("Phosphorus, Total (As P)", "Phosphorus, Total Orthophosphate (as P)", "Suspended Solids (Residue, Non-Filterable)")
analytes_plot3 <- setdiff(unique(filtered_df$analyte), c(analytes_plot1, analytes_plot2))

N_df    <- filtered_df %>% dplyr::filter(analyte %in% analytes_plot1)
P_df    <- filtered_df %>% dplyr::filter(analyte %in% analytes_plot2)
Misc_df <- filtered_df %>% dplyr::filter(analyte %in% analytes_plot3)
```

**CSV Export of selected site data saved as:** `r file`

# Events Captured
```{r}
unique_treatments <- filtered_df %>%
  dplyr::select(treatment.name, event_date) %>%
  dplyr::distinct() %>%
  dplyr::group_by(treatment.name) %>%
  dplyr::summarise(
    n_events = dplyr::n_distinct(event_date),
    collected = paste(sort(unique(format(event_date, "%Y-%m-%d"))), collapse = ", "),
    .groups = "drop"
  ) %>%
  dplyr::arrange(treatment.name)

DT::datatable(
  unique_treatments,
  caption = "Events Captured by Treatment: number of events and collection dates (YYYY-MM-DD)"
)
```

# Sample Standards
EPA analyte thresholds:

| Analyte                                      | Threshold | Units |
| :------------------------------------------- | :-------- | :---- |
| pH*                                          | 6.5–8.5   | —     |
| EC                                           |           |       |
| Phosphorus, total orthophosphate             |           |       |
| Phosphorus, total*                           | 0.0001    | mg/L  |
| Total dissolved solids**                     | 500       | mg/L  |
| Nitrogen, nitrate*                           | 10        | mg/L  |
| Nitrogen, nitrite*                           | 1         | mg/L  |
| Nitrogen, total kjeldahl                     |           |       |
| Suspended solids (residue, non-filterable)   |           |       |


*[EPA — National Primary Drinking Water Regulations](https://www.epa.gov/ground-water-and-drinking-water/national-primary-drinking-water-regulations)

**[EPA Nutrient Criteria Manual: Lakes & Reservoirs](https://www.epa.gov/sites/default/files/2018-10/documents/nutrient-criteria-manual-lakes-reservoirs.pdf)

CDPHE analyte thresholds:

| Analyte                                      | Threshold | Units |
| :------------------------------------------- | :-------- | :---- |
| pH (warm water)                              | 6.5–9     | —     |
| Nitrogen, nitrate                            | 10        | mg/L  |
| Nitrogen, nitrite                            | 1         | mg/L  |
| Total phosphorus (warm water)                | 0.17      | mg/L  |   
| Total phosphorus (cold water)                | 0.11      | mg/L  |

[CDPHE analyte thresholds](https://cdphe.colorado.gov/environment/water-quality/clean-water/water-quality-standards)  

# Exceedance Summary
```{r}
# Reg 31 exceedance table from filtered_df

# Standards dictionary 
standards <- tibble::tribble(
  ~analyte,      ~regulation,   ~threshold,  ~units,
  "pH",            "EPA",          8.5,      "None",    
  "pH",           "CDPHE",         9,        "None",  
  "NO3",           "EPA",          10,       "mg/L",
  "NO3",          "CDPHE",         10,       "mg/L",
  "NO2",           "EPA",          1,        "mg/L",  
  "NO2",          "CDPHE",         1,        "mg/L",
  "TKN",           "EPA",          NA_real_, "mg/L",
  "TKN",          "CDPHE",         NA_real_, "mg/L",
  "TP",            "EPA",          0.0001,   "mg/L", 
  "TP",           "CDPHE",         0.17,     "mg/L", 
  "TP",           "CDPHE",         0.11,     "mg/L",
  "OP",            "EPA",          NA_real_, "mg/L",
  "OP",           "CDPHE",         NA_real_, "mg/L", 
  "TDS",           "EPA",          500,      "mg/L",
  "TDS",          "CDPHE",         NA_real_, "mg/L",
  "TSS",           "EPA",          NA_real_, "mg/L",
  "TSS",          "CDPHE",         NA_real_, "mg/L",
  "EC",            "EPA",          NA_real_, "mg/L",
  "EC",           "CDPHE",         NA_real_, "mg/L", 
)

# 2) Helper: normalize analyte names to keys in 'standards'
map_analyte_to_key <- function(x) {
  x_low <- tolower(x)
  dplyr::case_when(
    str_detect(x_low, "^ph$") ~ "pH",
    str_detect(x_low, "nitrate") ~ "NO3",
    str_detect(x_low, "nitrite") ~ "NO2",
    str_detect(x_low, "total kjeldahl") ~ "TKN",
    str_detect(x_low, "total \\(?as p\\)?$") ~ "TP",
    str_detect(x_low, "total orthophosphate") ~ "OP",
    str_detect(x_low, "dissolved solids") ~ "TDS",
    str_detect(x_low, "suspended solids") ~ "TSS",
    str_detect(x_low, "specific conductance|^ec$") ~ "EC",
    TRUE ~ NA_character_
  )
}

# 3) Coalesce possible sample ID columns → 'sample_id'
sample_id_col <- coalesce(
  filtered_df$sample_id %>% suppressWarnings(),
  filtered_df$SampleID %>% suppressWarnings(),
  filtered_df$sample.id %>% suppressWarnings(),
  filtered_df$SAMPLE_ID %>% suppressWarnings(),
  filtered_df$Sample_ID %>% suppressWarnings()
)
filtered_df <- filtered_df %>%
  mutate(sample_id = sample_id_col)

# 4) Prepare data & join standards (NEW: join to LONG standards)
df_std_long <- filtered_df %>%
  mutate(
    analyte_key     = map_analyte_to_key(analyte),
    value           = suppressWarnings(as.numeric(result)),
    collected_date  = as.Date(collected)
  ) %>%
  left_join(
    standards,
    by = c("analyte_key" = "analyte"),
    relationship = "many-to-many"   # <- add this
  )


# A small edit before computing exceedance
df_std_long <- df_std_long %>%
  rename(std_units = units.y, sample_units = units.x)

# 5) Compute exceedances (long-format threshold check)
res_long <- df_std_long %>%
  mutate(
    exceeded = is.finite(threshold) & !is.na(value) & value > threshold,
    why = dplyr::if_else(
      exceeded,
      sprintf(
        "%s %s > %s %s (value=%.3g)",
        regulation, analyte_key,
        format(threshold, scientific = TRUE),
        std_units, value
      ),
      NA_character_
    )
  )

# Collapse back to one row per original sample/analyte (like before),
# combining EPA/CDPHE triggers into "EPA & CDPHE"
res <- res_long %>%
  group_by(across(setdiff(names(res_long), c("regulation","threshold","units","exceeded","why")))) %>%
  summarise(
    triggered = {
      regs <- regulation[exceeded %in% TRUE]
      if (length(regs) == 0) "None" else paste(sort(unique(regs)), collapse = " & ")
    },
    why = {
      msgs <- why[exceeded %in% TRUE & !is.na(why)]
      if (length(msgs) == 0) NA_character_ else paste(msgs, collapse = " | ")
    },
    .groups = "drop"
  )

# 6) Summary table (ONLY within-analyte percent)
n_total <- nrow(res)

summary_any <- res %>%
  mutate(exceeded_any = !is.na(triggered) & triggered != "None") %>%
  group_by(analyte_key) %>%
  summarise(
    count = sum(exceeded_any, na.rm = TRUE),
    pct_within_analyte = 100 * count / n(),
    .groups = "drop"
  ) %>%
  arrange(desc(count), analyte_key)

# -----------------------------
# Pretty summary table (DT)
# -----------------------------
summary_any_dt <- summary_any %>%
  mutate(
    pct_within_analyte = round(pct_within_analyte, 2)
  ) %>%
  select(analyte_key, count, pct_within_analyte)

dt_summary_any <- datatable(
  summary_any_dt,
  rownames = FALSE,
  extensions = c("Buttons", "FixedHeader"),
  options = list(
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    pageLength = 25,
    lengthMenu = list(c(10, 25, 50, 100),
                      c("10", "25", "50", "100")),
    autoWidth = FALSE,
    fixedHeader = TRUE,
    order = list(list(1, "desc")),
    class = "stripe hover compact",
    scrollX = TRUE,
    columnDefs = list(
      list(className = "dt-center", targets = "_all")
    ),
    headerCallback = JS(
      "function(thead, data, start, end, display) {",
      "  $(thead).find('th').css({'text-align': 'center'});",
      "}"
    )
  ),
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: center; font-weight: bold;",
    "Reg 31 exceedance summary by analyte"
  )
) %>%
  formatRound("pct_within_analyte", 2) %>%
  formatStyle("count", fontWeight = "bold")

dt_summary_any
```


# Interactive Box & Whisker Plot
```{r, warning=FALSE}
create_boxplot <- function(df, selected_location) {
  tryCatch({
    analytes <- unique(df$analyte.abbr)
    first_analyte <- analytes[1]
    df_filtered <- df[df$analyte.abbr == first_analyte, ]

    p <- ggplot(df_filtered,
                aes(x = treatment.name,
                    y = as.numeric(result),
                    fill = treatment.name)) +
      geom_boxplot(position = position_dodge(0), na.rm = TRUE) +
      geom_jitter(width = 0.2, size = 1.5, alpha = 0.6, na.rm = TRUE) +
      facet_grid(
        analyte.abbr ~ event_facet,
        scales = "free_y",
        space = "free_x",
        labeller = labeller(event_facet = event_date_only_labeller)
      ) +
      theme_minimal() +
      labs(
        title = paste("Box-and-Whisker plots for Location:", selected_location),
        x = "Event Type",
        y = "Result",
        fill = "Event Type"
      )

    ggplotly(p, tooltip = c("y"))
  }, error = function(e) {
    cat("An error occurred:", conditionMessage(e), "\n")
    NULL
  })
}

create_boxplot(filtered_df, loc)
```

```{r, fig.height=20, fig.width=12}
# dynamic x-axis in static plots
if (length(levels(as.factor(filtered_df$treatment.name))) > 1) {
  x_axis_var_name <- "treatment.name"
} else {
  x_axis_var_name <- "event.type"
}
```

# Static Bar Plots (Per Event)
```{r}
# ---- Put this right after you define `label_map` (once per report) ----
event_date_only_labeller <- function(values) {
  dates <- label_map$event_date[match(values, label_map$event_facet)]
  setNames(paste0("Event Date\n", format(dates, "%m-%d")), values)
}

# ---- Drop-in replacement for your plotting function ----
create_and_save_bar_plot_event <- function(df, output_dir, file_name, plot.title, x_axis_var) {
  tryCatch({
    if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

    df <- df %>%
      dplyr::mutate(
        result = suppressWarnings(as.numeric(result)),
        analyte = as.character(analyte),
        analyte.abbr = as.character(analyte.abbr),
        units = as.character(units)
      ) %>%
      dplyr::filter(is.finite(result))

    # Ensure "Inflow" appears first (if present)
    if (x_axis_var %in% names(df)) {
      df[[x_axis_var]] <- factor(
        df[[x_axis_var]],
        levels = c("Inflow", setdiff(unique(df[[x_axis_var]]), "Inflow"))
      )
    }

    # ---- Build facet row label: analyte on first line, units on second line
    df <- df %>%
      dplyr::mutate(
        analyte_name_for_plot = dplyr::if_else(
          !is.na(analyte.abbr) & analyte.abbr != "",
          analyte.abbr, analyte
        ),
        facet_var = paste0(analyte_name_for_plot, "\n(", units, ")")
      )

    # summary stats per event facet, analyte+units, and x-axis
    summary_df <- df %>%
      dplyr::group_by(.data[[x_axis_var]], facet_var, event_facet) %>%
      dplyr::summarise(
        mean = mean(result, na.rm = TRUE),
        sd   = sd(result, na.rm = TRUE),
        n    = dplyr::n(),
        .groups = "drop"
      )

    # facet ordering & labels
    summary_df$event_facet <- factor(summary_df$event_facet, levels = levels(df$event_facet))
    df$event_facet         <- factor(df$event_facet, levels = levels(df$event_facet))

    file_path_portrait  <- file.path(output_dir, file_name)
    file_path_landscape <- file.path(output_dir, paste0(tools::file_path_sans_ext(file_name), "_landscape.jpg"))

    bar_plot <- ggplot(summary_df, aes(x = .data[[x_axis_var]], y = mean, fill = .data[[x_axis_var]])) +
      geom_bar(stat = "identity", position = position_dodge(0), color = "black") +
      geom_errorbar(
        aes(ymin = pmax(mean - sd, 0), ymax = mean + sd),
        width = 0.2, position = position_dodge(0)
      ) +
      # raw points (jitter) in background
      geom_jitter(
        data = df,
        aes(x = .data[[x_axis_var]], y = result),
        width = 0.2, alpha = 0.6, size = 1,
        inherit.aes = FALSE,
        show.legend = FALSE
      ) +
      facet_grid(
        facet_var ~ event_facet,
        scales = "free_y",
        space = "free_x",
        labeller = labeller(event_facet = event_date_only_labeller) # date-only label
        # labeller = labeller(event_facet = event_labeller) # full label with event count
      ) +

      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        strip.text  = element_text(size = 12, lineheight = 1.05),
        legend.position = "bottom"
      ) +
      labs(
        title = plot.title,
        x = "Event Type",
        y = "Mean Result",
        fill = "Event Type"
      ) +
      scale_fill_brewer(palette = "Pastel1")

    ggsave(file_path_portrait,  bar_plot, width = 8.5, height = 11, units = "in")
    ggsave(file_path_landscape, bar_plot, width = 11, height = 8.5, units = "in")

    bar_plot
  }, error = function(e) {
    message("Error in create_and_save_bar_plot_event: ", e$message)
    NULL
  })
}


```

[N Plots](../site_selector_temporary_results/bar_N.jpg)
```{r}
bar_N    <- create_and_save_bar_plot_event(N_df, "./site_selector_temporary_results", "bar_N.jpg", "Forms of Nitrogen (per Event)", x_axis_var_name); bar_N
```
[P Plots](../site_selector_temporary_results/bar_P.jpg)
```{r}
bar_P    <- create_and_save_bar_plot_event(P_df, "./site_selector_temporary_results", "bar_P.jpg", "Forms of Phosphorus (per Event)", x_axis_var_name); bar_P
```

[Misc Plots](../site_selector_temporary_results/bar_Misc.jpg)
```{r}
bar_Misc <- create_and_save_bar_plot_event(Misc_df, "./site_selector_temporary_results", "bar_Misc.jpg", "Water Quality Analytes (per Event)", x_axis_var_name); bar_Misc
```

# Annual Means by Analyte & Treatment 
```{r}
create_facet_bar_plot_annual <- function(df, output_dir, file_name, x_axis_var, stratify_by_method = FALSE) {
  tryCatch({
    if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

    # Clean + coerce types once
    df <- df %>%
      dplyr::mutate(
        result       = suppressWarnings(as.numeric(result)),
        analyte      = as.character(analyte),
        analyte.abbr = as.character(analyte.abbr),
        units        = as.character(units),
        method.name  = as.character(method.name)
      ) %>%
      dplyr::filter(is.finite(result))

    # Ensure "Inflow" appears first (if present)
    if (x_axis_var %in% names(df)) {
      df[[x_axis_var]] <- factor(
        df[[x_axis_var]],
        levels = c("Inflow", setdiff(unique(df[[x_axis_var]]), "Inflow"))
      )
    }

    # ---- Build a single, reliable facet label: "Analyte (units)"
    # Prefer analyte.abbr if it exists, otherwise fall back to full analyte name.
    df <- df %>%
      dplyr::mutate(
        analyte_name_for_plot = dplyr::if_else(
          !is.na(analyte.abbr) & analyte.abbr != "",
          analyte.abbr,
          analyte
        ),
        facet_var = paste0(analyte_name_for_plot, " (", units, ")")
      )

    # Keep raw data version for point overlay
    raw_df <- df

    # Summaries (keep facet_var in the grouping so it cannot get “lost”)
    if (stratify_by_method) {
      summary_df <- df %>%
        dplyr::group_by(.data[[x_axis_var]], facet_var, method.name) %>%
        dplyr::summarise(
          mean = mean(result, na.rm = TRUE),
          sd   = sd(result, na.rm = TRUE),
          n    = dplyr::n(),
          .groups = "drop"
        )

      p <- ggplot(summary_df, aes(x = .data[[x_axis_var]], y = mean, fill = method.name)) +
        geom_bar(stat = "identity", position = position_dodge(width = 0.7), color = "black") +
        geom_errorbar(
          aes(ymin = pmax(mean - sd, 0), ymax = mean + sd),
          width = 0.2, position = position_dodge(width = 0.7)
        ) +
        # ---- RAW DATA POINTS (dodged by method) ----
        geom_point(
          data = raw_df,
          aes(x = .data[[x_axis_var]], y = result, color = method.name),
          position = position_jitterdodge(jitter.width = 0.15, dodge.width = 0.7),
          alpha = 0.06,
          size = 1.6,
          inherit.aes = FALSE,
          show.legend = FALSE
        )

    } else {
      summary_df <- df %>%
        dplyr::group_by(.data[[x_axis_var]], facet_var) %>%
        dplyr::summarise(
          mean = mean(result, na.rm = TRUE),
          sd   = sd(result, na.rm = TRUE),
          n    = dplyr::n(),
          .groups = "drop"
        )

      p <- ggplot(summary_df, aes(x = .data[[x_axis_var]], y = mean, fill = .data[[x_axis_var]])) +
        geom_bar(stat = "identity", position = "dodge", color = "black") +
        geom_errorbar(
          aes(ymin = pmax(mean - sd, 0), ymax = mean + sd),
          width = 0.2, position = position_dodge(width = 0.9)
        ) +
        # ---- RAW DATA POINTS ----
        geom_point(
          data = raw_df,
          aes(x = .data[[x_axis_var]], y = result),
          position = position_jitter(width = 0.15, height = 0),
          alpha = 0.06,
          size = 1.6,
          inherit.aes = FALSE,
          show.legend = FALSE
        )
    }

    p <- p +
      facet_wrap(~ facet_var, scales = "free_y") +
      theme_minimal() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text  = element_text(size = 11, face = "bold"),
        legend.position = "bottom"
      ) +
      labs(
        title = ifelse(
          stratify_by_method,
          "Annual Mean by Analyte (Stratified by Method)",
          "Annual Mean by Analyte"
        ),
        x = "Event Type",
        y = "Mean Result",
        fill = ifelse(stratify_by_method, "Method", x_axis_var)
      ) +
      scale_fill_brewer(palette = "Pastel1") +
      coord_cartesian(ylim = c(0, NA))

    file_path <- file.path(output_dir, file_name)
    ggsave(file_path, p, width = 10, height = 8, units = "in")

    p
  }, error = function(e) {
    message("Error in create_facet_bar_plot_annual: ", e$message)
    NULL
  })
}

# ---- calls (as requested) ----
facet_bar_without_method <- create_facet_bar_plot_annual(
  filtered_df,
  "./site_selector_temporary_results",
  "facet_bar_analytes_annual.jpg",
  x_axis_var_name,
  stratify_by_method = FALSE
); facet_bar_without_method

facet_bar_with_method <- create_facet_bar_plot_annual(
  filtered_df,
  "./site_selector_temporary_results",
  "facet_bar_analytes_method_annual.jpg",
  x_axis_var_name,
  stratify_by_method = TRUE
); facet_bar_with_method
```

# Raw Data
```{r}
DT::datatable(
  filtered_df %>%
    dplyr::mutate(collected = collected_disp) %>%
    dplyr::select(-collected_disp, -collected_local)
)
```

# Samples that Exceed Regulation 31 Standards
```{r}
reg31_exceedances <- res_long %>%
  # only actual violations (threshold exists AND exceeded == TRUE)
  filter(exceeded %in% TRUE) %>%
  transmute(
    sample_id,
    collected_date,
    analyte = analyte,          # original analyte label from filtered_df
    analyte_key,
    value,
    regulation,
    standard = sprintf("≤ %.3g", threshold),
    why
  ) %>%
  # make EPA/CDPHE into separate columns like before
  pivot_wider(
    names_from  = regulation,
    values_from = standard,
    names_glue  = "{regulation}_standard"
  ) %>%
  # bring back the combined trigger label ("EPA", "CDPHE", "EPA & CDPHE")
  left_join(
    res %>% select(sample_id, collected_date, analyte_key, value, triggered),
    by = c("sample_id", "collected_date", "analyte_key", "value")
  ) %>%
  # keep the columns in your original order
  transmute(
    sample_id,
    collected_date,
    analyte,
    value,
    EPA_standard = EPA_standard,
    CDPHE_standard = CDPHE_standard,
    triggered,
    why
  ) %>%
  arrange(collected_date, analyte, desc(value))

# -----------------------------
# Pretty exceedance samples table (DT)
# -----------------------------
reg31_exceedances_dt <- reg31_exceedances %>%
  mutate(
    value = round(value, 4)
  )

dt_reg31_exceedances <- datatable(
  reg31_exceedances_dt,
  rownames = FALSE,
  extensions = c("Buttons", "FixedHeader"),
  options = list(
    dom = "Bfrtip",
    buttons = c("copy", "csv", "excel"),
    pageLength = 25,
    autoWidth = TRUE,
    fixedHeader = TRUE,
    order = list(list(1, "asc")),  # collected_date
    class = "stripe hover compact",
    scrollX = TRUE
  ),
  filter = "top",
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-weight: bold;",
    "Reg 31 exceedance samples"
  )
) %>%
  formatStyle(
    "triggered",
    fontWeight = "bold"
  ) %>%
  formatStyle(
    "why",
    whiteSpace = "normal",
    lineHeight = "120%"
  )

dt_reg31_exceedances
```
