---
title: "Water Report"
author: "Caz Bell and A.J. Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
---


```{r, include=FALSE}
knitr::opts_chunk$set(toc = TRUE, echo = FALSE, warning = FALSE, message = FALSE)


```



# Introduction

This report details data collected from 2023-01-01 to `r Sys.Date()` by the Colorado State University (CSU) Agriculture Water Quality Program (AWQP). For more information, please contact AJ Brown at [Ansley.Brown\@colostate.edu](mailto:Ansley.Brown@colostate.edu){.email} or visit the CSU AWQP Website at <https://waterquality.colostate.edu>.



# Notes

-	All “0” values reported are non-detects (ND) reported by ALS
-	Total suspended solids (TSS), Electrical Conductivity (EC), and pH are performed by AWQP staff at CSU, whereas all other analytes are reported by ALS Laboratories
-	Methods used to derive each analyte can be found in the Raw Data table in the “method” column



```{r, include=FALSE}
package.list <- c("ggplot2",
                  "dplyr",
                  "DT",
                  "plotly"
)
packageLoad <- function(packages){
  for (i in packages) {
    if (!require(i, character.only = TRUE)) {
      install.packages(i)
      library(i, character.only = TRUE)
    }
  }
}
packageLoad(package.list)
```





```{r, include=FALSE}
#clean data source
source("file-merger.R", local = knitr::knit_global())
```


```{r, include=FALSE}
# Set directories
#report_path <- paste(dirname(getwd()), "/Example Report", sep = "")
data_path <- paste(dirname(getwd()), "/Data", sep = "")
print(data_path)
```

```{r, include=FALSE}
# import data
dat <- returnAllFiles(d = data_path, export = FALSE) 
#dat <- read.csv('all_files.csv', header = TRUE, stringsAsFactors = FALSE)
#make everything lowercase
colnames(dat) <- tolower(names(dat))
```




# Sites

```{r}
#create data table based on counts
counts <- dat %>% 
  group_by(location.name, analyte) %>% 
  summarise(count = n())
DT::datatable(counts, options = list(pageLength = length(counts$location.name)))
```


```{r}
make_scatter_plots <- function(dataframe) {
  
  dataframe <- ungroup(dataframe)
  # Get unique analytes and sort them in alphabetical order
  analytes <- sort(unique(dataframe$analyte))
  
  # Define custom colors for event types
  event_colors <- c("Inflow" = "#E69F00", "Outflow" = "#56B4E9", "Point Sample" = "#009E73")
  
  for (i in 1:length(analytes)) {
    # Subset data for the current analyte
    subset_analyte <- subset(dataframe, analyte == analytes[i])
    
    # Check if there are no results for the current analyte
     if (nrow(subset_analyte) == 0) {
      next  # Skip creating the plot for this analyte
    }
    
    # Get unique units for the current analyte
    units <- unique(subset_analyte$units)
  
      # Create scatter plot for the current analyte with custom colors
      plot <- ggplot(subset_analyte, aes(x =  format(collected, "%m/%d/%Y"), y = result, color = event.type)) +
        geom_point() + 
        scale_color_manual(values = event_colors) +  # Set custom colors for event types
        labs(x = "Date", y = paste0("Result (", units, ")"), title = paste(analytes[i], "Results")) +
        theme_bw()
      
      # Print the plot
      print(plot)
      
  
      
      # Add a new line after each plot
      cat("\n")
    }
  }


```



## ARDEC 2200

```{r}
#filter to site
ARDEC2200 <- dat %>% 
  filter(location.name == "ARDEC 2200")


# Call the scatter plot function for the site
make_scatter_plots(ARDEC2200)
```

## AVRC STAR

```{r}

AVRC <- dat %>% 
  filter(location.name == "AVRC STAR")

make_scatter_plots(AVRC)
```



## Barley

```{r}

Barley <- dat %>% 
  filter(location.name == "Barley")

make_scatter_plots(Barley)
```

## Below Stagecoach Dam

```{r}

BelowStage <- dat %>% 
  filter(location.name == "Below Stagecoach Dam")


# Call the scatter plot function with your dataframe
make_scatter_plots(BelowStage)
```



## Berthoud

```{r}
Berthoud <- dat %>% 
  filter(location.name == "Berthoud")


make_scatter_plots(Berthoud)
```


## Big Hollow

```{r}
BigHollow <- dat %>% 
  filter(location.name == "Big Hollow")


make_scatter_plots(BigHollow)
```


## Boulder Lake

```{r}
BoulderLake <- dat %>% 
  filter(location.name == "Boulder Lake")


make_scatter_plots(BoulderLake)
```



## Gunnison

```{r}
Gunnison <- dat %>% 
  filter(location.name == "Gunnison")


make_scatter_plots(Gunnison)
```

## Kerbel

```{r}
Kerbel <- dat %>% 
  filter(location.name == "Kerbel")


make_scatter_plots(Kerbel)
```

## Legacy

```{r}
Legacy <- dat %>% 
  filter(location.name == "Legacy")


make_scatter_plots(Legacy)
```

## Molina

```{r}
Molina <- dat %>% 
  filter(location.name == "Molina")


make_scatter_plots(Molina)
```


## Morrison Creek

```{r}
MorrisonCreek <- dat %>% 
  filter(location.name == "Morrison Creek")


make_scatter_plots(MorrisonCreek)
```

## Stage Coach Above

```{r}
SCA <- dat %>% 
  filter(location.name == "Stage Coach Above")


make_scatter_plots(SCA)
```

## Stage Coach In

```{r}
SCI <- dat %>% 
  filter(location.name == "Stage Coach In")


make_scatter_plots(SCI)
```

## Stagecoach

```{r}
Stagecoach <- dat %>% 
  filter(location.name == "Stagecoach")


make_scatter_plots(Stagecoach)
```

## Todds Ranch

```{r}
TR <- dat %>% 
  filter(location.name == "Todds Ranch")


make_scatter_plots(TR)
```

## Upper Yampa

```{r}
UY <- dat %>% 
  filter(location.name == "Upper Yampa")


make_scatter_plots(UY)
```


## Yellow Jacket

```{r}
YJ <- dat %>% 
  filter(location.name == "Yellow Jacket")


make_scatter_plots(YJ)
```



# Raw Data


```{r}
datatable(dat, colnames = c('Sample ID', 'Lab ID', 'Method', 'CAS Number', 'Analyte', 'Result', 'Units', 'Dilution', "Results Reported To", 'MDL', 'RL', 'Report Basis', 'Moisture (%)', 'Solid (%)', 'Non detect', 'Duplicate', 'Location', 'Treatment', 'Method Name', 'Event Type', 'Lab ID', 'Matrix', 'Collection Date', 'Recieved Date', 'Hold Time', 'Flag'))
```





