---
title: "water-report"
author: "AJ Brown"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This report details data collected from 2023-01-01 to `r Sys.Date()` by the Colorado State University (CSU) Agriculture Water Quality Program (AWQP). For more information, please contact AJ Brown at [Ansley.Brown\@colostate.edu](mailto:Ansley.Brown@colostate.edu){.email} or visit the CSU AWQP Website at <https://waterquality.colostate.edu>.

```{r, warning=FALSE, message=FALSE}
# import packages
package.list <- c("ggplot2",
                  "dplyr",
                  "DT",
                  "plotly"
                  )
packageLoad <- function(packages){
  for (i in packages) {
    if (!require(i, character.only = TRUE)) {
      install.packages(i)
      library(i, character.only = TRUE)
    }
  }
}
packageLoad(package.list)
#source("file-merger")

# Set directories
 # Working file_path
 # fix this to be relative path
file_path <- "../ALS-Data-Cleaning-Tool/all_files.csv/"
file_path <- "C:\\Users\\ansleybr\\Documents\\GitHub\\ALS-Data-Cleaning-Tool\\all_files.csv"
 # Set the report directory to the directory containing the selected file
directory <- paste(dirname(file_path),"/Example Report", sep = "")
print(directory)
# import data
# TODO: utilitze source() to import data from file-merge.R instead of reading 
# in the csv
dat <- read.csv(file_path, header = TRUE, stringsAsFactors = FALSE)
```

## Table of Contents

1.  [Site List](#sites)
2.  [Summary Plots](#plots)

## Site List <a name="sites"></a>

Here is a list of sites and their sample counts:

```{r}
# The DT package creates an interactive table for the HTML output.
# Here we create a summary table of location and test counts
counts <- dat %>% 
    group_by(location.name, ANALYTE) %>% 
    summarise(count = n())
DT::datatable(counts, options = list(pageLength = length(counts$location.name)))
```

## Summary Plots <a name="plots"></a>
```{r}
# Create a plot of the data using plotly
# group the data by "location", "test.type", and "value"
df_grouped <- dat %>%
    group_by(location.name, ANALYTE) %>%
    arrange(location.name)

# generate the button list dynamically based on unique "test.type" values
button_list <- lapply(unique(df_grouped$location.name), function(loc) {
  list(
    method = "update",
    args = list(
      list(visible = df_grouped$location.name == loc),
      list(title = paste("Test Type:",
        df_grouped$ANALYTE[1],
        "<br>Location:", loc)
        )
    ),
    label = loc
  )
}
)

# create the interactive box plots faceted by "ANALYTE"
p <- plot_ly(df_grouped, x = ~ANALYTE, y = ~RESULT, type = "box") %>%
  layout(title = "Water Analysis Results",
         xaxis = list(title = "Test Type"),
         yaxis = list(title = "Concentration (mg/L) or (ug/L; Selenium)"),
         updatemenus = list(
           list(
             y = 0.7,
             buttons = button_list
           )
         ),
         grid = list(rows = unique(df_grouped$ANALYTE))
        )

# show the plot
p
```